\Large\bf FlexibleSUSY --- A spectrum generator generator for supersymmetric models

### 2014-11-13 doc/flexiblesusy-paper.tex
Sort references in paper
clarify what is meant by "global sign" in the EWSB solutions

\fs assumes that each SUSY model contains Higgs bosons, which trigger
a spontaneous breaking of the electroweak symmetry.  The corresponding
EWSB consistency conditions are formulated in \fs at the one-loop
level as
%
\begin{align}
  0 &= \frac{\partial V^\text{tree}}{\partial v_i} - t_i,
  & & (i=1,\ldots,N)
  \label{eq:one-loop-ewsb-eq}
\end{align}
%
where $V^\text{tree}$ is the tree-level Higgs potential, $v_i$ is the
VEV corresponding to the Higgs fields $H_i$ and $t_i$ is the one-loop
tadpole diagram of $H_i$.  Already at the tree-level ($t_i = 0$) Eqs.\
\eqref{eq:one-loop-ewsb-eq} can have multiple solutions, depending on
which paramters are chosen to be fixed by these equations.  A
well-known example is the real MSSM, where the $\mu$-parameter is
chosen to be fixed by the EWSB equations.  In this model, the EWSB
equations, Eqs.\ \eqref{eq:one-loop-ewsb-eq} can only determine
$|\mu|$, while the sign of $\mu$ is not fixed.  This results in two
solutions for $\mu$ of the form
%
\begin{align}
  \mu = \sign\mu \cdot |\mu| .
  \label{eq:EWSB-solution-for-mu}
\end{align}
%
Currently, \fs handles only real model parameters and renormalizable
theories, which restricts the number of possible multiple solutions of
Eqs.\ \eqref{eq:one-loop-ewsb-eq} to be less or equal to $4N$.  In
some cases, as for example in the real MSSM, the different solutions
are related by one or more global signs as in
Eq.~\eqref{eq:EWSB-solution-for-mu}.  \fs recognizes such cases and
introduces these global signs as additional free parameters to allow
the user to chose between the different solutions.  The case of the
real MSSM Eq.~\eqref{eq:EWSB-solution-for-mu} is therefore handled
automatically in \fs, because $\sign\mu$ is automatically introduced
as an additional free parameter.

If, however, the different solutions are not related by global signs,
then \fs writes all solutions to the file
\code{models/<model>/<model>_tree_level_EWSB_solution.m}.  The user
has then the option to pick a particular solution by setting it in the
\code{TreeLevelEWSBSolution} variable in the model file.  This is for
example the case in the $Z_3$-violating NMSSM (\code{SMSSM}), where
$\mu$ is chosen to be fixed by the EWSB equations: There the
tree-level solution for $\mu$ has the form
%
\begin{align}
  \mu = -\frac{v_s \lambda}{\sqrt{2}} + \sign X \cdot \sqrt{\ldots}
\end{align}
%
with a non-global $\sign X$.  When \fs solves the tree-level EWSB
equations for $\mu$, $B\mu$ and $\xi_S$ it finds the two solutions and
writes them to \code{models/SMSSM/SMSSM_tree_level_EWSB_solution.m} in
the form
%
\begin{lstlisting}[language=Mathematica]
{
   {{B[\[Mu]] -> ...}},
   {
      {\[Mu] -> (-20*Sqrt[2]*vd*vS*(\[Lambda] + conj[\[Lambda]])
                 - Sqrt[...])/(80*vd)},
      {\[Mu] -> (-20*Sqrt[2]*vd*vS*(\[Lambda] + conj[\[Lambda]])
                 + Sqrt[...])/(80*vd)}
   },
   {{L[L1] -> ...}}
}
\end{lstlisting}
%
In the above solutions the dots stand for the full expression, which
is left out here for better readability.  Inspecting the two above
solutions for $\mu$, one finds that both can be parametrized by an
additional free sign in front of the \code{Sqrt[...]}.  The user can
now introduce an additional free sign by hand in the \code{MINPAR}
block
%
\begin{lstlisting}[language=Mathematica]
MINPAR = {
   {1, m0},
   {2, m12},
   {3, TanBeta},
   {4, Sign[X]},  (* <-- additional free sign *)
   {5, Azero}
};
\end{lstlisting}
%
and set the the tree-level solution, parametrized in terms of
\code{Sign[X]}, in the \code{TreeLevelEWSBSolution} variable in
the model file:
%
\begin{lstlisting}
TreeLevelEWSBSolution = {
   { B[\[Mu]], ... },
   { \[Mu], (-20*Sqrt[2]*vd*vS*(\[Lambda] + conj[\[Lambda]])
             + Sign[X] * Sqrt[...])/(80*vd) },
   { L[L1], ... }
};
\end{lstlisting}
%
One can now chose between the two solutions by setting entry $4$ in
the \code{MINPAR} block of the SLHA input file to either $+1$ or $-1$.
See \code{model_files/SMSSM/FlexibleSUSY.m.in} for a complete example
model file.  If the user decides to not pick a particular solution by
leaving the variable \code{TreeLevelEWSBSolution} empty, \fs tries to
find a solution to the tree-level EWSB equations numerically via an
iteration.  In this case, however, the user does not have the option
to chose between the different solutions.

@@

The spectrum generator shall accept
the low-energy neutrino masses and mixing angles
which determine \code{WOp}, the coefficients
of the $L H_u L H_u$ operator.
These $6 \times 6$ coefficients evolve to
the right-handed neutrino mass scale at which
they are matched to the neutrino Yukawa couplings \code{Yv} and
the right-handed neutrino masses \code{Mv}.
Since there are more degrees of freedom in the pair of
\code{Yv} and \code{Mv} than in \code{WOp},
one needs supplementary conditions in addition to the see-saw relation.
In the presented matching code, it is assumed that all mixing in \code{WOp}
stems from the left-handed rotation of \code{Yv} whose eigenvalues are
fixed to those specified by the user.
The non-trivial flavour structure of the neutrino Yukawa couplings then
causes running slepton mass matrices to acquire flavour-violating elements.
The output from the spectrum generator includes
the slepton mass matrices as well as
the resulting mass eigenvalues and mixing.
One might pass this outcome on to another routine
to calculate rates of lepton flavour violating processes.

>>>>

\fs assumes that each SUSY model contains Higgs bosons, which trigger
a spontaneous breaking of the electroweak symmetry.  The corresponding
EWSB consistency conditions are formulated in \fs at the one-loop
level as
%
\begin{align}
  0 &= \frac{\partial V^\text{tree}}{\partial v_i} - t_i,
  & & (i=1,\ldots,N)
  \label{eq:one-loop-ewsb-eq}
\end{align}
%
where $V^\text{tree}$ is the tree-level Higgs potential, $v_i$ is the
VEV corresponding to the Higgs fields $H_i$ and $t_i$ is the one-loop
tadpole diagram of $H_i$.  Already at the tree-level ($t_i = 0$) Eqs.\
\eqref{eq:one-loop-ewsb-eq} can have multiple solutions, depending on
which paramters are chosen to be fixed by these equations.  A
well-known example is the real MSSM, where the $\mu$-parameter is
chosen to be fixed by the EWSB equations.  In this model, the EWSB
equations, Eqs.\ \eqref{eq:one-loop-ewsb-eq} can only determine
$|\mu|$, while the sign of $\mu$ is not fixed.  This results in two
solutions for $\mu$ of the form
%
\begin{align}
  \mu = \sign\mu \cdot |\mu| .
  \label{eq:EWSB-solution-for-mu}
\end{align}
%
Currently, \fs handles only real model parameters and renormalizable
theories, which restricts the number of possible multiple solutions of
Eqs.\ \eqref{eq:one-loop-ewsb-eq} to be less or equal to $4N$.  In
some cases, as for example in the real MSSM, the different solutions
are related by one or more global signs for the parameters, as for example in
Eq.~\eqref{eq:EWSB-solution-for-mu}.  \fs recognizes such cases and
introduces these global signs as additional free parameters to allow
the user to chose between the different solutions.  The case of the
real MSSM Eq.~\eqref{eq:EWSB-solution-for-mu} is therefore handled
automatically in \fs, because $\sign\mu$ is automatically introduced
as an additional free parameter.

If, however, the different solutions are not related by global signs,
then \fs writes all solutions to the file
\code{models/<model>/<model>_tree_level_EWSB_solution.m}.  The user
has then the option to pick a particular solution by setting it in the
\code{TreeLevelEWSBSolution} variable in the model file.  This is for
example the case in the $Z_3$-violating NMSSM (\code{SMSSM}), which is shipped with \fs, where
$\mu$ is chosen to be fixed by the EWSB equations: There the
tree-level solution for $\mu$ has the form
%
\begin{align}
  \mu = -\frac{v_s \lambda}{\sqrt{2}} + \sign X \cdot \sqrt{\ldots}
\end{align}
%
with $\sign X$ as free parameter.  When \fs solves the tree-level EWSB
equations for $\mu$, $B\mu$ and $\xi_S$ it finds the two solutions and
writes them to \code{models/SMSSM/SMSSM_tree_level_EWSB_solution.m} in
the form
%
\begin{lstlisting}[language=Mathematica]
{
   {{B[\[Mu]] -> ...}},
   {
      {\[Mu] -> (-20*Sqrt[2]*vd*vS*(\[Lambda] + conj[\[Lambda]])
                 - Sqrt[...])/(80*vd)},
      {\[Mu] -> (-20*Sqrt[2]*vd*vS*(\[Lambda] + conj[\[Lambda]])
                 + Sqrt[...])/(80*vd)}
   },
   {{L[L1] -> ...}}
}
\end{lstlisting}
%
In the above solutions the dots stand for the full expression, which
is left out here for better readability.  Inspecting the two above
solutions for $\mu$, one finds that both can be parametrized by an
additional free sign in front of the \code{Sqrt[...]}.  The user can
now introduce an additional free sign by hand in the \code{MINPAR}
block
%
\begin{lstlisting}[language=Mathematica]
MINPAR = {
   {1, m0},
   {2, m12},
   {3, TanBeta},
   {4, Sign[X]},  (* <-- additional free sign *)
   {5, Azero}
};
\end{lstlisting}
%
and set the the tree-level solution, parametrized in terms of
\code{Sign[X]}, in the \code{TreeLevelEWSBSolution} variable in
the model file:
%
\begin{lstlisting}
TreeLevelEWSBSolution = {
   { B[\[Mu]], ... },
   { \[Mu], (-20*Sqrt[2]*vd*vS*(\[Lambda] + conj[\[Lambda]])
             + Sign[X] * Sqrt[...])/(80*vd) },
   { L[L1], ... }
};
\end{lstlisting}
%
One can now chose between the two solutions by setting entry $4$ in
the \code{MINPAR} block of the SLHA input file to either $+1$ or $-1$.
See \code{model_files/SMSSM/FlexibleSUSY.m.in} for a complete example
model file.  If the user decides to not pick a particular solution by
leaving the variable \code{TreeLevelEWSBSolution} empty, \fs tries to
find a solution to the tree-level EWSB equations numerically via an
iteration.  In this case, however, the user does not have the option
to chose between the different solutions.

@@

Since each model class has its own $\beta$-functions,
the spectrum generator contains two different sets of RGEs
that are connected by a matching object.
The program shall accept
the low-energy neutrino masses and mixing angles
which determine \code{WOp}, the coefficients
of the $L H_u L H_u$ operator.
These $6 \times 6$ coefficients evolve to
the right-handed neutrino mass scale at which
they are matched to the neutrino Yukawa couplings \code{Yv} and
the right-handed neutrino masses \code{Mv}.
Since there are more degrees of freedom in the pair of
\code{Yv} and \code{Mv} than in \code{WOp},
one needs supplementary conditions in addition to the see-saw relation.
In the presented matching code, it is assumed that all mixing in \code{WOp}
stems from the left-handed rotation of \code{Yv} whose eigenvalues are
fixed to those specified by the user.
The non-trivial flavour structure of the neutrino Yukawa couplings then
causes running slepton mass matrices to acquire flavour-violating elements.
The output from the spectrum generator includes
the slepton mass matrices as well as
the resulting mass eigenvalues and mixing.
One might pass this outcome on to another routine
to calculate rates of lepton flavour violating processes.
